The following list is the complete set of categories that are evaulated in the Syneval dataset, taken from \citet{linzen2018targeted} with slight alterations. There is a slight ambiguity in the category of negative polarity items, with two different types on constructions, one of sentences starting with \textit{most} and the other of sentences starting with \textit{most}, combined into one positive set; we choose to separate them into two different categories. For the full list of lexical items used to build variants of these constructions we refer the reader to \citet{linzen2018targeted}.

\begin{enumerate}[noitemsep]
  \item \textbf{Simple agreement}:
    \begin{enumerate}
      \item The farmer \textit{smiles}.
      \item *The farmer \textit{smile}.
    \end{enumerate}
  \item \textbf{Agreement in a sentential complement}:
    \begin{enumerate}
      \item The mechanics said the author \textit{laughs}.
      \item *The mechanics said the author \textit{laugh}.
    \end{enumerate}
  \item \textbf{Agreement in short VP coordination}:
    \begin{enumerate}
      \item The authors laugh and \textit{swim}.
      \item *The authors laugh and \textit{swims}.
    \end{enumerate}
  \item \textbf{Agreement in long VP coordination}:
    \begin{enumerate}
      \item The author knows many different foreign languages and \textit{enjoys} playing tennis with colleagues.
      \item *The author knows many different foreign languages and \textit{enjoy} playing tennis with colleagues.
    \end{enumerate}
  \item \textbf{Agreement across a prepositional phrase}:
    \begin{enumerate}
      \item The author next to the guards \textit{smiles}.
      \item *The author next to the guards \textit{smile}.
    \end{enumerate}
  \item \textbf{Agreement across a subject relative clause}:
    \begin{enumerate}
      \item The author that likes the security guards \textit{laughs}.
      \item *The author that likes the security guards \textit{laugh}.
    \end{enumerate}
  \item \textbf{Agreement across an object relative}:
    \begin{enumerate}
      \item The movies that the guard likes \textit{are} good.
      \item *The movies that the guard likes \textit{is} good.
    \end{enumerate}
  \item \textbf{Agreement across an object relative (no \textit{that})}:
    \begin{enumerate}
      \item The movies the guard likes \textit{are} good.
      \item *The movies the guard likes \textit{is} good.
    \end{enumerate}
  \item \textbf{Agreement in an object relative}:
    \begin{enumerate}
      \item The movies that the guard \textit{likes} are good.
      \item *The movies that the guard \textit{like} are good.
    \end{enumerate}
  \item \textbf{Agreement in an object relative (no \textit{that})}:
    \begin{enumerate}
      \item The movies the guard \textit{likes} are good.
      \item *The movies the guard \textit{like} are good.
    \end{enumerate}
  \item \textbf{Simple reflexive anaphora}:
    \begin{enumerate}
      \item The author injured \textit{himself}.
      \item *The author injured \textit{themselves}.
    \end{enumerate}
  \item \textbf{Reflexive in sentential complement}:
    \begin{enumerate}
      \item The mechanics said the author hurt \textit{himself}.
      \item *The mechanics said the author hurt \textit{themselves}.
    \end{enumerate}
  \item \textbf{Reflexive across a relative clause}:
    \begin{enumerate}
      \item The author that the guards like injured \textit{himself}.
      \item *The author that the guards like injured \textit{themselves}.
    \end{enumerate}
  \item \textbf{Simple NPI}:
    \begin{enumerate}
      \item \textit{No} authors have ever been famous.
      \item *\textit{Most} authors have ever been famous.
    \end{enumerate}
  \item \textbf{Simple NPI (the)}:
    \begin{enumerate}
      \item \textit{No} authors have ever been popular.
      \item *\textit{The} authors have ever been popular.
    \end{enumerate}
  \item \textbf{NPI across a relative clause}:
    \begin{enumerate}
      \item \textit{No} authors that \textit{the} guards like have ever been famous.
      \item *\textit{Most} authors that \textit{no} guards like have ever been famous.
    \end{enumerate}
  \item \textbf{NPI across a relative clause (the)}:
    \begin{enumerate}
      \item \textit{No} authors that \textit{the} guards like have ever been famous.
      \item *\textit{The} authors that \textit{no} guards like have ever been famous.
    \end{enumerate}
\end{enumerate}



% \begin{enumerate}
%   \item \textbf{Simple agreement}:  \\
%     The farmer \textit{smiles}/*\textit{smile}.
%   \item \textbf{Agreement in a sentential complement}:  \\
%     The mechanics said the author \textit{laughs}/*\textit{laugh}.
%   \item \textbf{Agreement in short VP coordination}:  \\
%     The authors laugh and \textit{swim}/*\textit{swims}.
%   \item \textbf{Agreement in long VP coordination}:  \\
%     The author knows many different foreign languages and \textit{enjoys}/*\textit{enjoy} playing tennis with colleagues.
%   \item \textbf{Agreement across a prepositional phrase}:  \\
%     The author next to the guards \textit{smiles}/*\textit{smile}.
%   \item \textbf{Agreement across a subject relative clause}:  \\
%     The author that likes the security guards \textit{laughs}/*\textit{laugh}.
%   \item \textbf{Agreement across an object relative}:  \\
%     The movies that the guard likes \textit{are}/*\textit{is} good.
%   \item \textbf{Agreement in an object relative}:  \\
%     The movies that the guard \textit{likes}/*\textit{like} are good.
%   \item \textbf{Simple reflexive anaphora}:  \\
%     The author injured \textit{himself}/*\textit{themselves}.
%   \item \textbf{Reflexive in sentential complement}:  \\
%     The mechanics said the author hurt \textit{himself}/*\textit{themselves}.
%   \item \textbf{Reflexive across a relative clause}:  \\
%     The author that the guards like injured \textit{himself}/*\textit{themselves}.
%   \item \textbf{Simple NPI}:  \\
%     \textit{No}/*\textit{most} authors have ever been famous.
%   \item \textbf{NPI across a relative clause}:  \\
%     No authors that the guards like have ever been famous.
%   \item \textbf{NPI across a relative clause (the)}:  \\
%     The authors that no guards like have ever been famous.
% \end{enumerate}
